---
title: "Q2"
author: "Arindom Baruah (32779267)"
format: html
---

# Exercise 2

A small cafe in the city of Melbourne is interested in determining whether the daily earnings depend on the weather. They compiled data for a period over 2000-2001 to study this question. Their data is provided. It has the following variables:

```{r}
#| echo: false
library(kableExtra)
library(tidyverse)
library(here)
desc <- tibble(var=c("dt", "wday", "revenue", "expend", "precip", "mint", "maxt", "source"),
               description=c("Date", 
              "Day of the week",
              "Daily revenue in hundreds, 11=1100", 
              "Daily expenses in hundreds",
              "Precipitation in mm", 
              "Minimum temperature, Celsius",
              "Maximum temperature, Celsius",
              "Source of the weather data"))
kbl(desc, table.attr = 'data-quarto-disable-processing="true"') |> 
  kable_styling(full_width = FALSE) |>
  column_spec(1, width="2cm", border_right = T) |>
  column_spec(2, width="10cm")
```

You need to conduct an **initial data analysis** to find and **report problems** that need to be fixed.


```{r}
cafe <- load("cafe.rda")
```


## Checking datatypes

The first step in the initial data analysis would be to check for the datatypes associated with each variable. We can identify the datatypes through the glimpse function as shown below.

```{r}
glimpse(cafe)
```



<div class="alert alert-block alert-warning">
We are able to observe the following issues with the datatypes of each variables:

1. <b> Date </b>: The date variable is in the character format and will need to be converted to datetime format.
2. <b> Expend </b>: The expend variable is in the character datatype while these are continuous numerical values. Hence, this variable will be converted to a numerical datatype.
3. <b> precip </b>: The precip variable which shows the precipitation value is a character variable while it contains continuous numerical values. Hence, this variable will be converted to a numerical variable.
</div>

