---
title: "Q3"
author: "Arindom Baruah (32779267)"
execute: 
  echo: true
  message: false
  warning: false
format: html
---



```{r load-library}
#| echo: false
library(kableExtra)
library(tidyverse)
library(here)
library(visdat)
library(lubridate)
library(plotly)
```

```{r load-data}

df_numbats <- read_csv("data/numbats.csv")

```
# Filter relevant data

In this section, we will filter our data in a manner such that we can focus on the sightings of Numbats across Australia from the years after 2010.

```{r}
df_numbats <- df_numbats %>% filter(year(eventDate) > 2010 )
```

# Count of daily sightings

The number of Numbat sightings for each event date are calculated through the code chunk below.

```{r}
df_numbats_daily <- df_numbats %>% group_by(eventDate) %>% summarise(daily_sightings = n())
```

Let us now visualise this data through a histogram as illustrated by @fig-hist

```{r}
#| label: fig-hist
#| fig-cap: "Histogram of daily sights of Numbats across Australia"

pl1 <- ggplot(data = df_numbats_daily,aes(x = daily_sightings)) + geom_histogram(fill = 'cornsilk',color = 'black') + labs(x = "Number of sightings per day",y = "Frequency of observations") + theme_classic()
pl1
```
<div class="alert alert-block alert-success">
💡 We can observe that majority of the daily sightings are around one per day. The distribution is right skewed with very daily sightings above 5 per day. 💡 </div>

While @fig-hist helped us visualise the data, we can obtain a similar distribution of the data using [stem and leaf plot](https://en.wikipedia.org/wiki/Stem-and-leaf_display). Let us create this visualisation using the stem function in R.

```{r}
stem(df_numbats_daily$daily_sightings)
```

<div class="alert alert-block alert-warning">
💡 As we can observe through the above visualisation, the stem and leaf plot illustrates a very similar understanding of the distribution of the data as the histogram plot. The stem leaf plot was able to capture the fact that majority of the observations were below 5 sightings per day and <strong> a right skew in the data </strong> could be observed. 💡 </div>
