---
title: "Q3"
author: "Arindom Baruah (32779267)"
execute: 
  echo: true
  message: false
  warning: false
format: html
---



```{r load-library}
#| echo: false
library(kableExtra)
library(tidyverse)
library(here)
library(visdat)
library(lubridate)
library(plotly)
```

```{r load-data}

df_numbats <- read_csv("data/numbats.csv")

```
# Filter relevant data

In this section, we will filter our data in a manner such that we can focus on the sightings of Numbats across Australia from the years after 2010.

```{r}
df_numbats <- df_numbats %>% filter(year(eventDate) > 2010 )
```

# Count of daily sightings

The number of Numbat sightings for each event date are calculated through the code chunk below.

```{r}
df_numbats_daily <- df_numbats %>% group_by(eventDate) %>% summarise(daily_sightings = n())
```


# Visualisation of daily sightings

Let us now visualise this data through a histogram as illustrated by @fig-hist

```{r}
#| label: fig-hist
#| fig-cap: "Histogram of daily sights of Numbats across Australia"

pl1 <- ggplot(data = df_numbats_daily,aes(x = daily_sightings)) + geom_histogram(fill = 'cornsilk',color = 'black') + labs(x = "Number of sightings per day",y = "Frequency of observations") + theme_classic()
pl1
```
<div class="alert alert-block alert-success">
💡 We can observe that majority of the daily sightings are around one per day. The distribution is right skewed with very daily sightings above 5 per day. 💡 </div>

While @fig-hist helped us visualise the data, we can obtain a similar distribution of the data using [stem and leaf plot](https://en.wikipedia.org/wiki/Stem-and-leaf_display). Let us create this visualisation using the stem function in R.

```{r}
stem(df_numbats_daily$daily_sightings)
```

<div class="alert alert-block alert-warning">
💡 As we can observe through the above visualisation, the stem and leaf plot illustrates a very similar understanding of the distribution of the data as the histogram plot. The stem leaf plot was able to capture the fact that majority of the observations were below 5 sightings per day and <strong> a right skew in the data </strong> could be observed. 💡 </div>


# Five number summaries based on Tukey's EDA

Tukey's EDA is based on the following 5 number summaries:

- Minimum value 
- Lower hinge 
- Median value  
- Upper hinge
- Maximum value

```{r}
#| label: tbl-fivnum
#| tbl-cap: "Five number summary based on Tukey's EDA" 

table <- tibble()
table <- tibble("Minimum value" = min(df_numbats_daily$daily_sightings),
                "Lower hinge" = quantile(df_numbats_daily$daily_sightings,0.25),
                "Median" = median(df_numbats_daily$daily_sightings),
                "Upper hinge" = quantile(df_numbats_daily$daily_sightings,0.75),
                "Maximum" = max(df_numbats_daily$daily_sightings))
table %>% kable()
```

@tbl-fivnum illustrates the five number summaries for the daily sightings of Numbats around Australia.

<div class="alert alert-block alert-success">
💡 The lower and upper hinge in Tukey's EDA correspond to the <b> 1st quartile (25th percentile) and 3rd quartile (75th percentile) </b> observations respectively. 💡 </div>


# Boxplot of daily sightings

Let us try to visualise the observations of daily sightings through a boxplot as illustrated by @fig-box

```{r}
#| label: fig-box
#| fig-cap: "Boxplot distribution of daily sightings"

pl4 <- ggplot(data = df_numbats_daily,aes(y = daily_sightings),color = 'black') + geom_boxplot() + labs(y = "Daily sightings") + theme_classic()
pl4
```

@fig-box illustrates a boxplot with the distribution of daily sightings of Numbats across Australia. We can observe a total of 4 outlier values which are illustrated as point observations in the boxplot. These are __those observations which are greater than the upper whisker__. The upper whisker is calculated as below.

$$\boxed{Upper Whisker = (1.5 * IQR) + Q3}$$
Based on the distribution of the histogram in @fig-hist and the stem plot, we can observe that the number of sightings which are above 5 per day are extremely low. This has been corroborated by the boxplot in @fig-box. As a result, we can conclude that the singular points marked in the boxplot are __reasonably marked as outliers based on the upper whisker value as calculated above__.
