---
title: "Q5"
author: "Arindom Baruah (32779267)"
execute: 
  echo: true
  message: false
  warning: false
format: html
number-sections: true
theme: journal
---


```{r}
#| echo: false
#| message: false
#| fig-width: 10
#| fig-height: 5
library(palmerpenguins)
library(colorspace)
library(patchwork)
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(nullabor)
library(broom)
```


# Data pre-processing

Let us check how the initial dataset looks like. @tbl-peng gives us a short glimpse of the dataset.

```{r}
#| label: tbl-peng
#| tbl-cap: "Penguin dataset from the Palmerpenguins package"
head(penguins) %>% kable %>% 
  kable_styling(bootstrap_options = c("bordered","hover"))
```

For further analysis, we will retain only the following variables and rename them as follows:

1. Species
2. Bill_length (Renamed as bl)
3. Bill_depth (Renamed as bd)
4. Flipper_length (Renamed as fl)
5. Body_mass_g (Renamed as bm)

We will additionally remove any missing values in the dataset.

```{r}

vars <- c("species","bill_length_mm",
          "bill_depth_mm","flipper_length_mm",
          "body_mass_g")
penguins_subset <- penguins %>% drop_na() %>% select(vars)
penguins_subset <- penguins_subset %>% rename("bl" = "bill_length_mm",
                                              "bd" = "bill_depth_mm",
                                              "fl" = "flipper_length_mm",
                                              "bm" = "body_mass_g")
```

The new pre-processed dataset is illustrated through @tbl-pengsub

```{r}
#| label: tbl-pengsub
#| tbl-cap: "Pre-processed dataset"

head(penguins_subset) %>% kable %>% 
  kable_styling(bootstrap_options = c("bordered","hover"))

```


# Plot visualisation of bm,fl and bl

```{r}
#| label: fig-charts
#| fig-cap: "Body mass Vs Flipper length Vs Bill length"
#| fig-subcap: 
#|   - "General plot for bm and fl"
#|   - "Variation across species"
#| layout-ncol: 2
pl1 <- ggplot(data = penguins_subset,aes(x = bm,y = fl)) + 
  geom_point() + labs(x = "Body mass (bm)",y = "Flipper length (fl)") + theme_classic() + theme(aspect.ratio=1)

pl2 <- ggplot(data = penguins_subset,aes(x = bm,y = bl)) + 
  geom_point(aes(color = species)) + labs(x = "Body mass (bm)",y = "Bill length (bl)",color = "Species") + 
  theme_classic() + scale_color_discrete_divergingx(palette="Zissou 1") 
pl1
pl2
```

The plots illustrated by @fig-charts show the relation between fl and bm across different species.

# Lineup of 20 plots 

## Plot A

Let us create the null plots for plot (a) in @fig-charts by permuting the residuals of the linear relationship between the var as shown by the code chunks below.

```{r}
#| label: fig-nulla
#| fig-cap: "Null plots for residuals of bm vs fl linear relationship"
set.seed(201349)
lmodel <- lm(formula = bm~fl,data =  penguins_subset)
ggplot(lineup(null_lm(bm~fl, method="rotate"),
              augment(lmodel)),
       aes(x=fl, y=.resid)) +
  geom_point() + geom_smooth(method = lm,color = 'blue',se=FALSE) +
  facet_wrap(~.sample, ncol=5) +
  theme_void() +
  theme(axis.text = element_blank(), 
        panel.border = element_rect(fill=NA, colour="black"))
```

@fig-nulla illustrates the null plots for data points of residuals __generated by randomising the residuals.__.

<div class="alert alert-block alert-warning">
💡

The null hypothesis and the alternative hypothesis are as follows:

$H_o$ = The variables body mass (bm) and flipper length (fl) have __a weak linear relationship__. \
$H_1$ = Body mass (bm) and flipper length(fl) do not have a weak linearly relationship to each other.

Method : Randomising the residuals using the null_lm function.
💡
</div>


## Plot B

Let us now create the null plots for the same variables fl and bm but add in a level of colors for different species as illustrated by @fig-nullb. We are going to __randomly permute the species of the penguins__ to test whether the variables of fl and bm have any correlation with species.

```{r}
#| label: fig-nullb
#| fig-cap: "Null plots for bm vs bl colored by different species"

set.seed(201344)
ggplot(lineup(null_permute("species"), penguins_subset, n=20),
       aes(x=bm, y=bl, colour=species)) +
  geom_point(alpha=0.7) +
  facet_wrap(~.sample, ncol=5) +
  colorspace::scale_color_discrete_divergingx(palette="Temps") +
  theme_bw() +
  theme(
    legend.position="none",
    axis.title = element_blank(),
    axis.text = element_blank()
  )
```


<div class="alert alert-block alert-warning">
💡

The null hypothesis and the alternative hypothesis are as follows:

$H_o$ = The variables body mass (bm) and flipper length (fl) have __no association with the type of penguin species__. \
$H_1$ = Body mass (bm) and flipper length(fl) are related are dependent upon the particular type of species.

Method : Permutation of the "species" variable using the null_permute function.
💡
</div>


# P-value calculation 

## Plot A

Based on the reports of 8 cases, following observations were recorded:

A total of 2 out of the 8 observations chose plot 18 as the most unique plot which is the actual plot.

Some of the reasons outlined for those who chose plot 18 are follows:

- Less noise in the data.
- Looks like a linear relationship.
- Better homoscedacity.

6 subjects however chose the remaining plots as there was no clear distinctive pattern in the scatterplot of the residuals.


Based on the above statistics, the P-value maybe calculated through the code-chunk below.


```{r pcalca}

p_val <- sum(dbinom(2:8,8,1/20))

```


<div class="alert alert-block alert-warning">
💡

Based on the statistics observed by the reports of the 8 test subjects, the P-value of 2 or more people choosing plot 18 as the most unique plot in the line up is __`r p_val`__. Such a value of score indicates that there is __not enough evidence to reject the null hypothesis__ in favor of the alternate hypothesis __which states that the variables of body mass (bm) and flipper length (fl) are weakly related to each other in a linear fashion__.
💡
</div>


## Plot B

Based on the reports of 8 cases, following observations were recorded:

All of the 8 observations recorded that they observed plot 10 to be the most unique plot in the lineup.

Some of the reasons outlined for those who chose plot 10 are as follows:

- Less dispersion in the residuals.
- Even distribution across the blue ideal line.
- A definitive pattern in the colors observed.
- Concentration of red color observed at the top end.




```{r pcalcb}

p_val <- sum(dbinom(8:8,8,1/20))

```


<div class="alert alert-block alert-warning">
💡

Based on the statistics observed by the reports of the 8 test subjects, the P-value of 8 choosing plot 10 as the most unique plot in the line up is __`r p_val`__. Such a low score indicates that there is overwhelming evidence to reject the null hypothesis in favor of the alternate hypothesis __which states that the variables of body mass (bm) and flipper length (fl) are indeed related to the type of penguin species__.
💡
</div>

# Resources

The above analysis was undertaken with the help of the following software and packages:

1. __RStudio__: Integrated Development for R. RStudio, PBC, Boston, MA URL http://www.rstudio.com/.
2. __ggplot2__: H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New York, 2016.
3. __tidyverse__: Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K,
  Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019). “Welcome to the tidyverse.” _Journal of Open Source Software_,
  *4*(43), 1686. doi:10.21105/joss.01686 <https://doi.org/10.21105/joss.01686>.
4. __knitr__: Yihui Xie (2014) knitr: A Comprehensive Tool for Reproducible Research in R. In Victoria Stodden, Friedrich Leisch and Roger D. Peng, editors, Implementing
  Reproducible Computational Research. Chapman and Hall/CRC. ISBN 978-1466561595.
5. __rmarkdown__: Allaire J, Xie Y, Dervieux C, McPherson J, Luraschi J, Ushey K, Atkins A, Wickham H, Cheng J, Chang W, Iannone R (2023). _rmarkdown: Dynamic Documents for R_.
  R package version 2.23, <https://github.com/rstudio/rmarkdown>.
6. __kableExtra__: Zhu H (2023). _kableExtra: Construct Complex Table with 'kable' and Pipe Syntax_. http://haozhu233.github.io/kableExtra
7. __visdat__: Tierney N (2017). “visdat: Visualising Whole Data Frames.” _JOSS_, *2*(16), 355. doi:10.21105/joss.00355 <https://doi.org/10.21105/joss.00355>,
  <http://dx.doi.org/10.21105/joss.00355>.
